---
title: "Intro"
execute: 
  echo: false
  message: false
  warning: false
---

Some factors can predict Alzheimer's Disease with more accuracy than others. People with low Functional Assessment test scores, low Activities of Daily Living scores, memory complaints, and behavioral problems had the highest odds of predicting Alzheimer's. Using health data from 2149 patients ages 60-90, this project explores the most effective predictors of Alzheimer's disease as of now. It seems that physical functioning, rather than the expected cognitive functioning or age, has a greater odd of predicting Alzheimer's disease in a person age 60-90.





```{r}
#|label: setup

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(ggplot2))

```


```{r}
#|cache: true

x <- read_csv("alzheimers_disease_data.csv", show_col_types = FALSE)

x <- x %>%
  mutate(
    ADL = if_else(ADL <= 5, "lowADL", "highADL"),    
    Age = if_else(Age < 75, "below75", "above75"),
    FunctionalAssessment = if_else(FunctionalAssessment < 5, "lowFA", "highFA")
  )

model <- brm(
  formula = Diagnosis ~ Age + Gender + FunctionalAssessment + ADL + MemoryComplaints + BehavioralProblems,
  data = x,
  family = bernoulli(),
  refresh = 0,
  silent = 2,
  seed = 9
)

modelf <- fixef(model)
modelf2 <- as.data.frame(modelf) %>%
  rownames_to_column(var = "term") %>%
  arrange(desc(abs(Estimate))) %>%
  mutate(OddsRatio = exp(Estimate))

ggplot(modelf2, aes(x = reorder(term, -abs(Estimate)), y = OddsRatio)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_text(aes(label = round(OddsRatio, 2)), vjust = -0.5) +
  labs(
    title = "Odds Ratios of Model Coefficients",
    x = "Predictor",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

library(tidyverse)
library(brms)
library(ggplot2)

x <- read_csv("alzheimers_disease_data.csv")

x <- x %>%
  mutate(
    ADL = if_else(ADL <= 5, "lowADL", "highADL"),    
    Age = if_else(Age < 75, "below75", "above75"),
    FunctionalAssessment = if_else(FunctionalAssessment < 5, "lowFA", "highFA")
  )

model <- brm(
  formula = Diagnosis ~ Age + Gender + FunctionalAssessment + ADL + MemoryComplaints + BehavioralProblems,
  data = x,
  family = bernoulli(),
  refresh = 0,
  silent = 2,
  seed = 9
)

modelf <- fixef(model)
modelf2 <- as.data.frame(modelf) %>%
  rownames_to_column(var = "term") %>%
  arrange(desc(abs(Estimate))) %>%
  mutate(OddsRatio = exp(Estimate),
         CI_lower = exp(Q2.5),
         CI_upper = exp(Q97.5),
         label = sprintf("OR: %.2f\n95%% CI: %.2f - %.2f", OddsRatio, CI_lower, CI_upper))

ggplot(modelf2, aes(x = reorder(term, Estimate), y = OddsRatio)) +
  geom_point(size = 3, color = "blue") +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2, color = "blue") +
  coord_flip() +
  labs(
    title = "Forest Plot of Model Coefficients",
    x = "Predictor",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = label, y = OddsRatio), hjust = -0.3, size = 3) +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = "Odds Ratios with 95% CI"))
```

